---
title: "2021 EFA steps"
author: "Tim Marchand"
date: "5/13/2021"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacman)
```

## Introduction

This document describes the step by step process of conducting
exploratory factor analysis from POSgrams of the BNC and CMC corpora.

The steps are as follows:

1.  Preliminaries

    1.  Load libraries

    2.  Load data set

    3.  Create functions

2.  Exploratory factor analysis

    1.  Preliminary steps

    2.  Step 1

    3.  Step 2

    4.  Step 3

    5.  Summary

3.  Internal consistency reliability

4.  Results presentation

## Preliminaries

### Load libraries
```{r}
p_load(tidyverse, psych, tidylog)
```

### Load data
We load the data from the outputs of the udpipe and wmatrix codes for comaprison.

```{r}
##import the file from POS project
dat <- read_csv(here::here("efa","data","efa_table.csv"))

## create matrix for efa
master <- dat %>% select(-c(2)) %>% column_to_rownames(var = "file")
```

A quick look at the data:
```{r}
head(dat) 
dim(master)
```



### Create functions
The following functions will be useful when looking at the EFA output from the "pysch" package.

```{r}
`%notin%` <- Negate(`%in%`) # always useful


# specify the cut point and get the loadings in a tidy format
get_loadings <- function(x, cut = 0) {
  #get sorted loadings
  loadings <- fa.sort(x)$loadings %>% round(3)
  #supress loadings
  loadings[loadings < abs(cut)] <- ""
  #get additional info
  add_info <- cbind(x$communalities, 
                    x$uniquenesses,
                    x$complexity) %>%
    # make it a data frame
    as.data.frame() %>%
    # column names
    rename("Communality" = V1,
           "Uniqueness" = V2,
           "Complexity" = V3) %>%
    #get the item names from the vector
    rownames_to_column("POSgram")
  #build table
  loadings %>%
    unclass() %>%
    as.data.frame() %>%
    rownames_to_column("POSgram") %>%
    left_join(add_info) %>%
    mutate(across(where(is.numeric), round, 3))
}

## get the factor correlations table
get_fa_cors <- function(x){
  x %>% pluck("Phi")
}

## get various fit statistics
get_BIC <- function(x){
  x %>% pluck("BIC")
}

get_RMSEA <- function(x){
  x %>% pluck("RMSEA")
}

get_TLI<- function(x){
  x %>% pluck("TLI")
}

get_cfi <- function(x){
  stat <- x %>% pluck("STATISTIC")
  dof <- x %>% pluck("dof")
  null.chisq <- x %>% pluck("null.chisq")
  null.dof <- x %>% pluck("null.dof")
  
  cfi <- (stat-dof)/(null.chisq - null.dof)
  return(cfi)
}
```

